### 2018-04-03 architecture.tex
split the huge file in smaller ones.

%!TEX root=restructured.tex

\section{Architecture}

REMEMBER THE GOAL: Our main goal with the \tool issues to allow the integration of the dashboard with an API with minimal effort.

\ins{The first important question that such a 
service monitoring system should provide is 
information about what endpoints are being
used and by whom}.

TODO: Find references, arguments, related work
that provide more details about why understanding
who uses endpoints, and which are used are important.

This is clearly the case in static analysis (see paper
by haenni and lungu...) but should even more be the
case in APIs.

  \newpage

  % This used to be in the first section... 
  % Maybe it fits better localized here: 
  Data collected by the wrappers are persisted in a local database. The SQLAlchemy Object Relational Mapper\footnote{\url{https://www.sqlalchemy.org/}} on top of SQLite\footnote{\url{https://www.sqlite.org/}} is used for this purpose.

TODO: TALK ABOUT THE METAMODEL...

    \begin{figure}[ht!]
      \centering
        \includegraphics[width=0.8\linewidth]{db_schema}
        \caption{The model that supports the views presented in this paper}
        \label{fig:sep}
    \end{figure}

  As discussed in the introductory section, in previous work~\cite{vogel2017low} we introduced \tool, a drop-in Python library that allows developers to monitor their Flask-based Python web applications with minimal effort.
%
  The \tool is implemented for Python 3.6 and is available on the Python Package Index repository\footnote{\url{https://pypi.python.org/pypi/flask-monitoring-dashboard/1.8}} from where it can be installed by running \install from the command line. 
%  
  The source code of the \tool is published under a permissive MIT license and is available on GitHub\footnote{\url{https://github.com/flask-dashboard}}.