### 2018-04-05 ease of use.tex
cleaning up...

ease of use:

statsd
collectd

Graphite used to be fairly difficult to install.

https://www.digitalocean.com/community/tutorials/how-to-install-and-use-graphite-on-an-ubuntu-14-04-server

25 pages document

https://community.netapp.com/t5/OnCommand-Storage-Management-Software-Articles-and-Resources/How-to-install-Graphite-and-Grafana/ta-p/109456

recipe for installing: 
-- tested recipe on RHEL 6 & 7, and Ubuntu, to get you up and running fast

how to install graphite:

https://www.digitalocean.com/community/tutorials/how-to-install-and-use-graphite-on-an-ubuntu-14-04-server

immense...

how to install collectd: 
https://www.digitalocean.com/community/tutorials/how-to-configure-collectd-to-gather-system-metrics-for-graphite-on-ubuntu-14-04

immense...