### 2018-04-09 overhead-table.tex
nicer and nicer ...

%!TEX root=paper.tex

\newcommand{\yes}{\checkmark}

\begin{table}[tb]

	\centering

	\begin{tabular}{rllrr}
		\toprule
		\bfseries Endpoint & \bfseries Dash. & \bfseries All Out. & \bfseries Mean (ms) & \bfseries SD (ms) \\

		\midrule

	         available\_languages  &    &	    &   6.3 &  2.3 \\ 
	         available\_languages  &  \yes &    &  19.9 &  5.6 \\ 
	         available\_languages  &  \yes &  \yes &  29.4 &  5.4 \\ \\

	                user\_article  &    &	    &  13.9 &  2.6 \\
	                user\_article  &  \yes &    &  26.9 &  3.2 \\
	                user\_article  &  \yes &  \yes &  44.8 & 10.0 \\ \\

	    create\_default\_ex...     &    &	    &  276.0 & 22.4 \\ 
	    create\_default\_ex...     &  \yes &    & 313.6 & 43.2 \\
	    create\_default\_ex...     &  \yes &  \yes & 385.3 & 16.9 \\

		\bottomrule

	\end{tabular}
	\caption{Response times for three endpoints run 500 times in three different conditions each. Dash. = with dashboard. All Out. = every call being considered an outlier}
	\label{tab:benchmark}

\end{table}